%& -shell-escape
%\documentclass[a4paper]{aa}
\documentclass[printer]{aa}
% \documentclass{article}
\usepackage[pagebackref,colorlinks,citecolor=blue,urlcolor=blue,linkcolor=blue]{hyperref}
\usepackage[varg]{txfonts}
\usepackage{graphicx}	
\usepackage{natbib}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{tabularx}
\usepackage{hhline}
\usepackage{makecell}
\usepackage{placeins}
\usepackage{tikz}
\usepackage{subcaption}
\usepackage{lipsum}
\usetikzlibrary{arrows,positioning,shapes} %,optics}
\usepackage{xspace}
\usepackage{mathtools}
\usepackage{{booktabs}}

%\usepackage[squaren,Gray]{SIunits}
\usepackage{siunitx}
%\usepackage{units} % nicefrac
%\newcommand{\units}[1]{$\mathrm{#1}$}

%instrument 
\newcommand{\SD}{StarDICE\xspace}
\newcommand{\Oldthorlabs}{SM05PD1B\xspace}
\newcommand{\Newthorlabs}{SM05PD3A\xspace}
\newcommand{\Qsolar}{Q_{\mathrm{solar}}}
\newcommand{\Qsolarmes}{Q_{\mathrm{solar}}^{\mathrm{mes}}}
\newcommand{\Qsolarcal}{Q_{\mathrm{solar}}^{\mathrm{cal}}}

\newcommand{\Qphot}{Q_{\mathrm{phot}}}
\newcommand{\Qphotmes}{Q_{\mathrm{phot}}^{\mathrm{mes}}}
\newcommand{\Qphotcal}{Q_{\mathrm{phot}}^{\mathrm{cal}}}
\newcommand{\Qphotcont}{Q_{\mathrm{phot}}^{\mathrm{532}}}

\newcommand{\Qccd}{Q_{\mathrm{ccd}}}
\newcommand{\Qccdcal}{Q_{\mathrm{ccd}}^{\mathrm{cal}}}
\newcommand{\Qccdmes}{Q_{\mathrm{ccd}}^{\mathrm{mes}}}
\newcommand{\Qccdcont}{Q_{\mathrm{ccd}}^{\mathrm{532}}}

\newcommand{\Qspectro}{Q_{\mathrm{spectro}}}
\newcommand{\Qspectrocont}{Q_{\mathrm{spectro}}^{\mathrm{532}}}
\newcommand{\Qspectromain}{Q_{\mathrm{spectro}}^{\mathrm{main}}}

\newcommand{\Ssolarstat}{\sigma_{\mathrm{solar}}^{\mathrm{stat}}}
\newcommand{\Sphotstat}{\sigma_{\mathrm{phot}}^{\mathrm{stat}}}
\newcommand{\Qghost}{Q_{\mathrm{ghost}}}
\newcommand{\Rspectro}{R_{\mathrm{spectro}}}
\newcommand{\Rcbp}{R_{\mathrm{CBP}}}
\newcommand{\Rtel}{R_{\mathrm{tel}}}

\newcommand{\Esolar}{\epsilon_{\mathrm{SC}}}
\newcommand{\Espectro}{\epsilon_{\mathrm{spectro}}}
\newcommand{\Espectrocont}{\epsilon_{\mathrm{spectro}}^{\mathrm{532}}}
\newcommand{\Ephot}{\epsilon_{\mathrm{phot}}}
\newcommand{\Ephotcont}{\epsilon_{\mathrm{phot}}^{\mathrm{532}}}
\newcommand{\Eccd}{\epsilon_{\mathrm{ccd}}(\lambda)}

\newcommand{\spinhole}{\SI{75}{\micro\meter}\xspace}
\newcommand{\bpinhole}{\SI{5}{\milli\meter}\xspace}

\newcommand{\mes}{{}_{\mathrm{, \, mes}}}
\newcommand{\cont}{{}_{\mathrm{, \, 532}}}  %cont plutôt que 532 ?
\newcommand{\contcomp}{{}_{\mathrm{, \, comp}}}  %cont plutôt que 532 ?

\newcommand{\Kghost}{K_{G_1/G_0}(\lambda)}
\newcommand{\Kpinholes}{K_\mathrm{\spinhole/\bpinhole}(\lambda)}

\newcommand{\Rwindow}{R_{\mathrm{win}}(\lambda)\xspace}
\newcommand{\Rccd}{R_{\mathrm{ccd}}(\lambda)\xspace}

\newcommand{\bkg}{\mathrm{bkg}}
\newcommand{\Kghostfit}{K_{G/A}}
\newcommand{\Kghostfitfirst}{K_{G_1/A}}

%\newcommand{\Kg_fit}{\mathrm{K_{G_{(1, 2)}/A}}}

\newcommand{\todo}[1]{\textbf{\textcolor{red}{[#1]}}\xspace}
\newcommand{\com}[1]{\textbf{\textcolor{orange}{(#1)}}\xspace}
\newcommand{\up}[1]{\textsuperscript{#1}}
\renewcommand\tabularxcolumn[1]{m{#1}}


\usepackage[outputdir={fig},debug]{dot2texi}
%\usepackage{draftwatermark}
\usepackage{xspace}
%\SetWatermarkText{DRAFT}
%\SetWatermarkScale{6}
%\SetWatermarkLightness{0.9}
\newcommand{\angexp}{\circ{a}}
\newcommand{\texp}{\ensuremath{\tau}\xspace}
\newcommand{\FixMe}[1]{\textbf{\textcolor{red}{[#1]}}\xspace}
\newcommand{\SZF}[1]{\textbf{\textcolor{blue}{[#1]}}\xspace}
\newcommand{\MARC}[1]{\textbf{\textcolor{green}{[#1]}}\xspace}
\newcommand{\THIERRY}[1]{\textbf{\textcolor{green}{[#1]}}\xspace}

\newcolumntype{M}[1]{>{\centering\arraybackslash}m{#1}}
%\renewcommand{\arraystretch}{2}
%\setlength{\arrayrulewidth}{1.25pt}

\graphicspath{{fig/}}
\makeatletter
\newcommand*\ExpandableInput[1]{\@@input#1 }
\makeatother
\DeclareMathOperator*{\med}{med}


\title{StarDICE III: Characterization of the photometric instrument with a Collimated Beam Projector}

\input{./sections/authors.tex}
\abstract{The measurement of Type Ia supernovae magnitudes provides cosmological distances, which can be used to constrain dark energy parameters. Yet, current and upcoming large photometric surveys require a substantial improvement in the calibration precision of their photometry to reduce systematic uncertainties in cosmological constraints.}{The StarDICE experiment is designed to establish accurate broadband flux references for these surveys, aiming for sub-percent precision in magnitude measurements. This requires a precise measurement of the filter bandpasses of both the StarDICE and survey instruments with sub-nanometer accuracy. To that end, we have developed the Collimated Beam Projector (CBP), an optical device capable of calibrating the throughput of an astronomical telescope and of its filters.}{The CBP is built from a tunable laser source and a reversed telescope to emit a parallel monochromatic light beam that is continuously monitored in flux and wavelength. The CBP output light flux is measured using a large area photodiode, previously calibrated relative to a NIST photodiode. We then derive the \SD{} telescope throughput and filter transmissions from the CBP measurements, anchoring it to the absolute calibration provided by the NIST.}{After carefully analyzing the systematic uncertainties, we have achieved sub-nanometer accuracy in determining filter central wavelengths, measured each filter transmission with a precision of \textasciitilde 0.5\% per \SI{1}{nm} bin, and detected out-of-band leakages at the \num{e-4} relative value. Furthermore, we have synthesized the equivalent transmission for full pupil illumination from four sample positions in the StarDICE telescope mirror, with an accuracy of approximately \SI{0.2}{nm} for central wavelengths and \SI{7}{mmag} for broadband fluxes.}{We have demonstrated our ability to characterize a telescope throughput down to the millimagnitude, and paved the way for future developments, such as a portable CBP version for in-situ transmission monitoring.}


\authorrunning{T. Souverin et al.}
\titlerunning{Characterisation of the photometric instrument with a Collimated Beam Projector}

\begin{document}

\authorrunning{T. Souverin et al.}
\titlerunning{Characterisation of the photometric instrument with a Collimated Beam Projector}


\maketitle

%\section{Currently happening}
%
%\begin{itemize}
%\item SC linearity
%\item pinhole choice
%\item dark current caracterization
%\item Photo-diode change
%\end{itemize}

%\tableofcontents

\input{sections/1_introduction}

\input{sections/2_laboratory_setup}

\input{sections/3_cbp_response}

%\input{sections/4_instrument_model}
%\clearpage

\input{sections/4_stardice_response}

\input{sections/5_pupil_stitching}

\input{sections/6_discussion}



\bibliographystyle{aa}
\bibliography{cbp}

\begin{acknowledgements}
  This work received support from the Programme National Cosmology et
  Galaxies (PNCG) of CNRS/INSU with INP and IN2P3, co-funded by CEA
  and CNES and from the DIM ACAV program of the Île-de-France region.
  CWS, EU, and SB are grateful to the US Department of Energy for support under Cosmic Frontier award DE- SC0007881. 

  \input{sections/contributions}
  
  \input{sections/standard}
  
\end{acknowledgements}

\appendix
\input{sections/8_appendix}




\end{document}
