This paper has undergone internal review in the LSST Dark Energy Science Collaboration. The internal reviewers were Johan Bregeon and Parker Fagrelius.\\

Thierry Souverin is the primary author of the paper, and he contributed to the hardware assembly, the data taking, the data analysis and the paper edition. Jérémy Neveu is the primary author's advisor and participated to the hardware assembly, the data taking, the data analysis and the paper edition.  Marc Betoule is the StarDICE PI and participated to the hardware assembly, data taking, data reduction, lead the pupill stitching analysis and participated to the paper edition. Sébastien Bongard contributed to the logistics, the hardware assembly, the data analysis and the paper edition.  Christopher W. Stubbs is the instigator of the CBP concept and he contributed to a large part of the hardware funding, and to the solar cell calibration. Elana Urbach participated to the hardware assembly, the data taking and the solar cell calibration. 
Sasha Brownsberger participated to the hardware assembly, the data taking, and the solar cell calibration. Pierre Éric Blanc is member of the StarDICE collaboration. Johann Cohen Tanugi is member of the StarDICE collaboration. Sylvie Dagoret-Campagne is member of the StarDICE collaboration. Fabrice Feinstein is member of the StarDICE collaboration. Delphine Hardin is member of the StarDICE collaboration. Claire Juramy is member of the StarDICE collaboration. Laurent Le Guillou is member of the StarDICE collaboration and contributed with software and hardware support. Auguste Le Van Suu is member of the StarDICE collaboration. Marc Moniez is member of the StarDICE collaboration. Éric Nuss (posthumous) is member of the StarDICE collaboration. Bertrand
Plez is member of the StarDICE collaboration. Nicolas Regnault is member of the StarDICE collaboration. Eduardo Sepulveda is member of the StarDICE collaboration and contributed with software and hardware support. Kélian Sommer is member of the StarDICE collaboration. All members of the StarDICE collaboration contributed by their general support to the collaboration and to the collaboration review of the paper.

